---
title: "Travel Survey Unweighted Summary"
author: "suzanne"
date: "2023-01-29"
output:
  html_document: default
  pdf_document: default
---

These summaries are a quick quality control check on the unweighted survey data. We will compare across 2017, 2019, 2021, and 2023 where feasible. The purpose is to identify major problems that need to be fixed in the data, not to fully validate the data since it's not weighted yet.

```{r include=FALSE, message=FALSE, warning=FALSE}
library(DBI)
library(dplyr)
library(data.table)
library(ggplot2)
library(tidyr)
library(janitor)
library(scales)
source('unweighted-summary-functions.R')

```

### 2023 data is structured differently; get the data and then bind with 2017, 2019, 2021
```{r }
library(data.table) # general data manipulation
library(srvyr)      # survey statistics
library(dplyr)      # helpers for srvyr (mutate, group_by, and %>% pipe), 
                    # plus function near(), used in qa/qc checks 
library(stringr)    # string helper functions
library(flextable)  # formatted tables
library(ggplot2)    # plotting
library(scales)     # ggplot2 helpers
library(plotly)     # interactive plots
library(rstudioapi) # for setting working directory
library(psrcelmer)
library(tidyr)

# Point to where the 2023 codebook lives:
codebook_path = 'J:/Projects/Surveys/HHTravel/Survey2023/Data/codebook/PSRC_Combined_Codebook_2023_groupings.xlsx'

# Read Data  ===================================================================



hh_rulesy<-get_table(db_name = "hhts_cleaning", "HHSurvey", "Household")
person_rulesy<-get_table(db_name = "hhts_cleaning", "HHSurvey", "Person")
day_rulesy<-get_table(db_name = "hhts_cleaning", "HHSurvey", "Day")
trip_raw<-get_table(db_name = "HouseholdTravelSurvey2023", "combined_data", "v_trip")%>%filter(hh_day_iscomplete_b==1)
trip_rulesy<-get_table(db_name = "hhts_cleaning", "HHSurvey", "Trip")

## Codebook -------------------------------------------------------------------- 

variables = readxl::read_xlsx(
  codebook_path,
  sheet = 'variable_list_2023'
)

values =  readxl::read_xlsx(
  codebook_path,
  sheet = 'value_labels_2023'
)

setDT(variables)
setDT(values)
```



## Count records on households, persons, day, trip tables
##Number of Households Surveyed
```{r}
hh_count<-get_hhts_no_weights(survey=c("2017", "2019", "2021"), level='h', vars='household_id')%>%
          group_by(survey_year)%>%
          count()

hh_count_23<-hh_rulesy%>%count()%>%
  mutate(survey_year=2023)
# Household Count
hh_count<-rbind(hh_count,hh_count_23)
hh_count
```
```{r}


hh<-get_hhts_no_weights(survey=c("2019", "2021"), level='h', vars='cityofseattle')

hh_23<-lookup_names(hh_rulesy, '2023', 'hhid', 'cityofseattle' )%>%mutate(survey_year= '2023')


hh<-rbind(hh, hh_23)

hh_summ<-category_totals_wide(hh, 'cityofseattle')

hh_summ

```

```{r}


#hh<-get_hhts_no_weights(survey=c("2019", "2021"), level='h', vars='cityofbellevue')

hh_23<-lookup_names(hh_rulesy, '2023', 'hhid', 'cityofbellevue' )%>%mutate(survey_year= '2023')


#hh<-rbind(hh, hh_23)

hh_summ<-category_totals_wide(hh_23, 'cityofbellevue')

hh_summ

```
```{r}


hh<-get_hhts_no_weights(survey=c("2019", "2021"), level='h', vars='hhincome_detailed')

hh_23<-lookup_names(hh_rulesy, '2023', 'hhid', 'hhincome_detailed' )%>%mutate(survey_year= '2023')


hh<-rbind(hh, hh_23)

hh_summ<-category_totals_wide(hh, 'hhincome_detailed')

hh_summ

hh_shares<-category_shares_wide(hh, 'hhincome_detailed')

```
```{r}


hh<-get_hhts_no_weights(survey=c("2019", "2021"), level='h', vars='hhsize')

hh_23<-lookup_names(hh_rulesy, '2023', 'hhid', 'hhsize' )%>%mutate(survey_year= '2023')


hh<-rbind(hh, hh_23)

hh_summ<-category_totals_wide(hh, 'hhsize')

hh_summ

hh_shares<-category_shares_wide(hh, 'hhsize')

```
```{r}


hh<-get_hhts_no_weights(survey=c("2019", "2021"), level='h', vars='sample_segment')

hh_23<-lookup_names(hh_rulesy, '2023', 'hhid', 'sample_segment' )%>%mutate(survey_year= '2023')


hh<-rbind(hh, hh_23)

hh_summ<-category_totals_wide(hh, 'sample_segment')

hh_summ

hh_shares<-category_shares_wide(hh, 'sample_segment')

```

##Number of People Surveyed
```{r}
person_count<-get_hhts_no_weights(survey=c("2017", "2019", "2021"), level='p', vars='person_id')%>%
          group_by(survey_year)%>%
          count()



person_count_23<-person_rulesy%>%count()%>%
  mutate(survey_year=2023)

person_count<-rbind(person_count,person_count_23)
person_count
```

##Number of Observed Days
```{r}
day_count<-get_hhts_no_weights(survey=c("2017", "2019", "2021"), level='d', vars='day_id')%>%
          group_by(survey_year)%>%
          count()
day_count_23<-day_rulesy%>%count()%>%
  mutate(survey_year=2023)

day_count<-rbind(day_count,day_count_23)
day_count
#Day Count
```


```{r}


trip_count<-get_hhts_no_weights(survey=c("2017", "2019", "2021"), level='t', vars='trip_id')%>%
          group_by(survey_year)%>%
          count()

trip_count_23<-trip_rulesy%>%ungroup()%>%count()%>%
  mutate(survey_year='2023_cleaned')

trip_count_raw<-trip_raw%>%ungroup()%>%count%>%
  mutate(survey_year='2023_raw')

trip_count<-trip_count%>%mutate(survey_year=as.character(survey_year))

trip_count<-rbind(trip_count,trip_count_raw, trip_count_23)
trip_count

```


## Household level
### prev_res_factors_forced
```{r }
hh_forced<-get_hhts_no_weights(survey=c("2019", "2021"), level='h', vars='prev_res_factors_forced')

hh_forced_23<-lookup_names(hh_rulesy, '2023', 'hhid', 'prev_res_factors_forced' )%>%mutate(survey_year= '2023')


hh_forced<-rbind(hh_forced, hh_forced_23)

tbl_counts<-category_shares(hh_forced, 'prev_res_factors_forced')

```

```{r}
ggplot(tbl_counts, aes(x=as.factor(survey_year), y=share, fill=prev_res_factors_forced))+geom_bar(stat = "identity")
```


```{r}


workplace<-get_hhts_no_weights(survey=c("2019", "2021"), level='p', vars='workplace')

workplace_23<-lookup_names(person_rulesy, '2023', 'hhid', 'workplace' )%>%mutate(survey_year= '2023')


workplace<-rbind(workplace, workplace_23)

tbl_counts<-category_shares(workplace, 'workplace')


```

## Persons level
### workplace

```{r }
person_workplace<-get_hhts_no_weights(survey=c("2017", "2019", "2021"), level='p', vars='workplace')
tbl_counts<-category_shares(person_workplace, 'workplace')

```


```{r}

ggplot(tbl_counts, aes(x=as.factor(survey_year), y=share, fill=workplace))+
  geom_bar(stat = "identity")

```
## Trips by purpose


```{r}


purpose<-get_hhts_no_weights(survey=c("2019", "2021"), level='t', vars='dest_purpose')

purpose_23<-lookup_names(trip_rulesy, '2023', 'tripid', 'dest_purpose' )%>%mutate(survey_year= '2023')


purpose<-rbind(purpose, purpose_23)

tbl_counts<-category_shares(purpose, 'dest_purpose')


```

#
```{r }
ggplot(tbl_counts, aes(x=survey_year, y=share, fill=dest_purpose))+
  geom_bar(stat = "identity")

```
## Trips by mode

```{r}


mode<-get_hhts_no_weights(survey=c("2019", "2021"), level='t', vars='mode_1')

mode_23<-lookup_names(trip_rulesy, '2023', 'tripid', 'mode_1' )%>%mutate(survey_year= '2023')


mode<-rbind(mode, mode_23)


tbl_counts<-category_shares(mode, 'mode_1')

#ggplot(tbl_counts, aes(x=survey_year, y=share, fill=mode_1))+
#  geom_bar(stat = "identity")
```
## Trips by distance

```{r}
distance<-get_hhts_no_weights(survey=c("2019", "2021"), level='t', vars='trip_path_distance')%>%rename(distance_miles=trip_path_distance)%>%mutate(distance_miles=as.numeric(distance_miles))

distance_23<-trip_rulesy%>%select(distance_miles)%>%mutate(survey_year= '2023')%>%mutate(distance_miles=as.numeric(distance_miles))


distance<-rbind(distance, distance_23)

median_dist<-distance%>%group_by(survey_year)%>%
  summarise(median_dist=median(distance_miles, na.rm=TRUE))
median_dist


```
```{r}
distance<-distance%>%filter(distance_miles<200)
ggplot(distance, aes(x=distance_miles))+geom_histogram(aes(fill=survey_year), position="dodge", alpha=0.4)+facet_wrap(vars(survey_year))
```


# Trips by purpose and mode

# Trips by mode and distance
