---
title: "Travel Survey Unweighted Summary"
author: "suzanne"
date: "2023-06-21"
output:
  html_document: default
  pdf_document: default
---

These summaries are a quick quality control check on the unweighted survey data. We will compare across 2017, 2019, 2021, and 2023 where feasible. The purpose is to identify major problems that need to be fixed in the data, not to fully validate the data since it's not weighted yet.

```{r include=FALSE, message=FALSE, warning=FALSE}
library(DBI)
library(dplyr)
library(data.table)
library(ggplot2)
library(tidyr)
library(janitor)
library(scales)
source('unweighted-summary-functions.R')

```

## Count records on households, persons, day, trip tables

```{r}
hh_count<-get_hhts_no_weights(survey=c("2017", "2019", "2021"), level='h', vars='household_id')%>%
          group_by(survey_year)%>%
          count()
# Household Count
hh_count

person_count<-get_hhts_no_weights(survey=c("2017", "2019", "2021"), level='p', vars='person_id')%>%
          group_by(survey_year)%>%
          count()

# Person Count
person_count

day_count<-get_hhts_no_weights(survey=c("2017", "2019", "2021"), level='d', vars='day_id')%>%
          group_by(survey_year)%>%
          count()

#Day Count
day_count

trip_count<-get_hhts_no_weights(survey=c("2017", "2019", "2021"), level='t', vars='trip_id')%>%
          group_by(survey_year)%>%
          count()

#Trip Count
trip_count

```


## Household level
### prev_res_factors_forced
```{r }
hh_forced<-get_hhts_no_weights(survey=c("2019", "2021"), level='h', vars='prev_res_factors_forced')
tbl_counts<-category_shares(hh_forced, 'prev_res_factors_forced')

```

```{r}
ggplot(tbl_counts, aes(x=as.factor(survey_year), y=n, fill=prev_res_factors_forced))+
  geom_bar(stat = "identity")
ggplot(tbl_counts, aes(x=as.factor(survey_year), y=share, fill=prev_res_factors_forced))+
  geom_bar(stat = "identity")
```

## Persons level
### workplace

```{r }
person_workplace<-get_hhts_no_weights(survey=c("2017", "2019", "2021"), level='p', vars='workplace')
tbl_counts<-category_shares(person_workplace, 'workplace')

```


```{r}
ggplot(tbl_counts, aes(x=as.factor(survey_year), y=n, fill=workplace))+
  geom_bar(stat = "identity")
ggplot(tbl_counts, aes(x=as.factor(survey_year), y=share, fill=workplace))+
  geom_bar(stat = "identity")

```

### student status

```{r }
person_student<-get_hhts_no_weights(survey=c("2017", "2019", "2021"), level='p', vars='student')
tbl_counts<-category_shares(person_student, 'student')

```

### industry
```{r }
person_industry<-get_hhts_no_weights(survey=c( "2021"), level='p', vars='industry')
tbl_counts<-category_shares(person_industry, 'industry')

```



### race and ethnicity
#### todo? many categories


## Trip summaries

### trips per person per day

```{r }
trips_purpose<-get_hhts_no_weights(survey=c("2017", "2019", "2021"), level='t', vars='dest_purpose_cat')
tbl_counts<-category_shares(trips_purpose, 'dest_purpose_cat')

```


```{r}
tbl_wide_tot<- tbl_counts%>%select(survey_year, n, dest_purpose_cat)%>%
    pivot_wider(names_from= survey_year, values_from=n)
tbl_wide_tot
```

```{r}

trip_counts<-trips_purpose%>%
    group_by(survey_year, dest_purpose_cat)%>%
    summarise(n=n())
  
people_count<-get_hhts_no_weights(survey=c("2017", "2019", "2021"), level='p', vars='person_id')%>%
  group_by(survey_year)%>%count()

```

```{r}
day_trips<-day_count %>% 
              left_join(trip_counts, by=join_by(survey_year), suffix=c(".day", ".trip"))%>%
              mutate(trips_person=n.trip/n.day)


trips_purp_year<-day_trips%>%pivot_wider(names_from=survey_year, values_from=trips_person, id_cols=dest_purpose_cat)%>%adorn_totals("row")


trips_purp_year

```
###Trips per person by purpose for people under 18

```{r}
trips_purpose_kids<-get_hhts_no_weights(survey=c("2017", "2019", "2021"), level='t', vars=c('dest_purpose_cat', 'age_category'))%>%filter(age_category=='Under 18 years')%>%
  group_by(survey_year, dest_purpose_cat)%>%count()


kids_day_count<-get_hhts_no_weights(survey=c("2017", "2019", "2021"), level='d', vars=c('day_id', 'age_category'))%>%filter(age_category=='Under 18 years')%>%group_by(survey_year)%>%count()
```
### Kids trip summary
```{r}
day_trip_kids<-kids_day_count %>% 
              left_join(trips_purpose_kids, by=join_by(survey_year), suffix=c(".day", ".trip"))%>%
              mutate(trips_person=n.trip/n.day)


kids_trips_purp_year<-day_trip_kids%>%pivot_wider(names_from=survey_year, values_from=trips_person, id_cols=dest_purpose_cat)%>%adorn_totals("row")


kids_trips_purp_year

```
−	Mean and median trip distances overall and by broad purposes.
```{r}
trip_lengths<-get_hhts_no_weights(survey=c("2017", "2019", "2021"), level='t', vars=c('dest_purpose_cat', 'trip_path_distance'))

# missing trip lengths
for(year in c("2017", "2019", "2021")){
  trips_lengths_yr<-trip_lengths%>%filter(survey_year==!!year)
  print(paste(year, 'Trip Length Summary'))
  print(summary(trips_lengths_yr$trip_path_distance))
}

# Median Trip Lengths over time and purpose
trip_purp<-trip_lengths%>%
group_by(survey_year, dest_purpose_cat)%>%
drop_na()%>% 
summarise(trip_path_dist_median=median(trip_path_distance))%>%
  pivot_wider(names_from=survey_year, values_from=trip_path_dist_median, id_cols=dest_purpose_cat)

```
### Transit Trips
## Transit Trips by Access Mode
```{r}
trip_mode<-get_hhts_no_weights(survey=c("2017", "2019", "2021"), level='t', vars=c('main_mode'))
mode_summary<-category_shares(trip_mode, 'main_mode')
```
```{r}
transit_access_mode<-get_hhts_no_weights(survey=c("2017", "2019", "2021"), level='t', vars=c('mode_acc'))
mode_acc_summary<-category_shares(transit_access_mode, 'mode_acc')

```
