---
title: Summarizing Rulesy Error Codes
  Fixie
author: "PSRC"
date: "2023-10-04"
output:
  pdf_document:
    toc: yes
    toc_depth: '4'
  html_document:
    toc: yes
    toc_depth: 4
    toc_float: yes
    css: styles.css
---

```{r }
library(data.table) # general data manipulation
library(srvyr)      # survey statistics
library(dplyr)      # helpers for srvyr (mutate, group_by, and %>% pipe), 
                    # plus function near(), used in qa/qc checks 
library(stringr)    # string helper functions
library(flextable)  # formatted tables
library(ggplot2)    # plotting
library(scales)     # ggplot2 helpers
library(plotly)     # interactive plots
library(rstudioapi) # for setting working directory
library(psrcelmer)
library(tidyr)

# Point to where the 2023 codebook lives:
codebook_path = 'J:/Projects/Surveys/HHTravel/Survey2023/Data/data_deliverable_81823/codebook_guide/PSRC_Combined_Codebook_2023_08162023.xlsx'

# Read Data  ===================================================================



hh_rulesy<-get_table(db_name = "HouseholdTravelSurvey2023", "combined_data", "v_household")
person_rulesy<-get_table(db_name = "HouseholdTravelSurvey2023", "combined_data", "v_person")
day_rulesy<-get_table(db_name = "HouseholdTravelSurvey2023", "combined_data", "v_day")
trip_raw<-get_table(db_name = "HouseholdTravelSurvey2023", "combined_data", "v_trip")
trip_rulesy_before<-get_table(db_name = "hhts_cleaning", "HHSurvey", "Trip")
trip_error_flags<-get_table(db_name = "hhts_cleaning", "HHSurvey", "trip_error_flags")
error_types<-get_table(db_name = "hhts_cleaning", "HHSurvey", "error_types")





## Codebook -------------------------------------------------------------------- 

variables = readxl::read_xlsx(
  codebook_path,
  sheet = 'variable_list_2023'
)

values =  readxl::read_xlsx(
  codebook_path,
  sheet = 'value_labels_2023'
)

setDT(variables)
setDT(values)
```

Create visualizable data tables, before and after the changes. Also show the errors
```{r}
trip_rulesy<-trip_rulesy_before%>%mutate(tripid=as.character(tripid))
trip_rulesy<-left_join(trip_rulesy, trip_error_flags, by = c('person_id', 'tripnum'), suffix=c('_rulesy', '_error'))%>%group_by(person_id, tripnum)%>%filter(row_number()==1)


trip_raw<-trip_raw%>%mutate(person_id=as.character(person_id))
person_raw<-person_rulesy%>%mutate(person_id=as.character(person_id))
trip_raw<-left_join(trip_raw, person_raw, by ='person_id')

trip_rulesy_subset<-trip_rulesy%>%
  select(error_flag, person_id, day_id, tripnum, tripid, depart_time_timestamp, arrival_time_timestamp, mode_1, distance_miles, travel_time, speed_mph, origin_lat, origin_lng, dest_lat, dest_lng, origin_purpose,dest_purpose,depart_time_timestamp,arrival_time_timestamp, arrive_date, arrival_time_hour, arrival_time_minute, depart_time_hour, depart_time_minute)%>%arrange(person_id, depart_time_timestamp, arrival_time_timestamp)%>%head(500)

trip_raw<-trip_raw%>%mutate(tripid=as.character(tripid))
trip_rulesy<-trip_rulesy%>%mutate(person_id=as.character(person_id))
trip_rulesy<-left_join(trip_rulesy, person_raw, by ='person_id')%>%
  select(error_flag, daynum,person_id, day_id, tripnum, tripid, depart_time_timestamp, arrival_time_timestamp, mode_1, distance_miles, travel_time, speed_mph, origin_lat, origin_lng, dest_lat, dest_lng, origin_purpose,dest_purpose,depart_time_timestamp,arrival_time_timestamp, arrive_date, arrival_time_hour, arrival_time_minute, depart_time_hour, depart_time_minute, user_added, user_merged, user_split, diary_platform, trace_quality_flag, hh_iscomplete_a, hh_iscomplete_b)%>%arrange(person_id, depart_time_timestamp, arrival_time_timestamp)

 # trip_rulesy_raw<-left_join(trip_raw, trip_rulesy, by='tripid', suffix=c('_before', '_after'))%>%
 #   select(error_flag,person_id_before, daynum_before,day_id_before, day_id_after, tripnum_after, tripid,origin_purpose_before, origin_purpose_after, dest_purpose_before, dest_purpose_after, arrive_date_before, arrive_date_after,arrival_time_hour_before,arrival_time_minute_before,depart_time_hour_before, depart_time_minute_before, depart_time_timestamp, arrival_time_timestamp,depart_time_minute_after,origin_lng_before, origin_lat_before,dest_lng_before,dest_lng_after, dest_lat_before,dest_lat_after, mode_1_before, mode_1_after, trace_quality_flag_before, day_is_complete_a_before, day_is_complete_b_before)%>%arrange(person_id_before, arrive_date_before, arrival_time_hour_before, arrival_time_minute_before, depart_time_hour_before, depart_time_minute_before)
```


```{r }
# first find the person ids of  trips with the error code

errors_edits<-trip_rulesy%>%group_by(user_added, user_merged, user_split,diary_platform, trace_quality_flag, hh_iscomplete_a, error_flag)%>%count()
```


```{r }
trips_edits<-trip_rulesy%>%mutate(user_edits_rmove=
                                    ifelse((diary_platform=='browser' | diary_platform=='call') |( user_added!=1 & user_merged!=1 & user_split!=1),'no edit', 'edit' ))


```


```{r}
errors_edits_no_user_edits<-trips_edits%>%group_by(user_edits_rmove, error_flag)%>%count()
write.csv(errors_edits_no_user_edits, 'errors_edits_no_user_edits.csv')
```


```{r }
# first find the person ids of  trips with the error code

errors_edits_no_user_edits_complete<-trips_edits%>% group_by(hh_iscomplete_b, user_edits_rmove, error_flag)%>%count()

write.csv(errors_edits_no_user_edits_complete, 'errors_edits_complete_b_no_user_edits.csv')
```



```{r}

errors_edits_complete_b<-trips_edits%>%group_by(hh_iscomplete_b, error_flag)%>%count()
write.csv(errors_edits_complete_b, 'errors_edits_hh_is_complete_b.csv')
```


```{r }
# first find the person ids of  trips with the error code

errors_edits_no_user_edits_complete<-trips_edits%>% group_by(hh_iscomplete_b, user_edits_rmove, error_flag)%>%count()

write.csv(errors_edits_no_user_edits_complete, 'errors_edits_complete_b_no_user_edits.csv')
```

```{r}
errors_complete_b<-trip_rulesy%>%filter(hh_iscomplete_b==1 & error_flag=='mode_1 missing')

mode_1_missing_trips<-errors_complete_b%>%filter(error_flag=='mode_1 missing')
persons_days_group<-mode_1_missing_trips%>%select(person_id,daynum)%>%distinct()
```

```{r }
mode_1_missing_days<-merge(persons_days_group, trip_rulesy, by=c('person_id','daynum'))%>%arrange(person_id, daynum, depart_time_timestamp, arrival_time_timestamp)
```

```{r}
label_values<-function(df, varname, values, new_name){
  vals<- values%>%filter(variable==sym(varname))
  df<-df%>%
      left_join(vals, by=join_by(!!ensym(varname)==value))%>%
      rename(!!ensym(new_name):=final_label)
  
  return(df)
}
mode_1_missing_purpose<-mode_1_missing_days%>%group_by(dest_purpose)%>%count()

mode_1_missing_purpose<-label_values(mode_1_missing_purpose, 'dest_purpose', values, 'destination purpose')


mode_1_missing_mode<-mode_1_missing_days%>%group_by(mode_1)%>%count()

mode_1_missing_mode<-label_values(mode_1_missing_mode, 'mode_1', values, 'destination purpose')

```

`

```{r}
sample_problem_days<-read.csv('bad_recs_is_complete_a.csv')
```

```{r}
days_detail<-left_join(sample_problem_days, trip_rulesy, by=c('person_id', 'daynum'))%>%select(error_flag, person_id, day_id, tripnum, tripid, depart_time_timestamp, arrival_time_timestamp, mode_1, distance_miles, travel_time, speed_mph, origin_lat, origin_lng, dest_lat, dest_lng, origin_purpose,dest_purpose,depart_time_timestamp,arrival_time_timestamp, arrive_date,trace_quality_flag,day_is_complete_a,user_added, user_merged, user_split)
```

```{r}
write.csv(days_detail, 'flagged_days_is_complete_a.csv')
```


### function to summarize a variable across flags
```{r}
var_error_compare<-function(df, dfname, id, var_name){
  #id=!!ensym(id)
  #var_name=!!ensym(var_name)
  df<-df%>%filter(!is.na(error_flag))
  df_cols<-df%>%select(!!ensym(id), !!ensym(var_name), error_flag)
  vals<- values%>%filter(variable==sym(var_name))
  df_cols<-df_cols%>%
    left_join(vals, by=join_by(!!ensym(var_name)==value))
  df_summary<-df_cols%>%
              group_by(final_label, error_flag)%>%
              count()
  df_summary<-df_summary%>%
              rename(!!quo_name(dfname):=n)
  
}
```


```{r}

trip_trace_quality_data_rulesy<-var_error_compare(df=trip_rulesy,dfname='trip_rulesy',id='tripid', var_name='trace_quality_flag')

```

### Error flags by mode
There are very high numbers of walk, transit trips with the change mode purpose and/or missing next trip link or missing prior trip link.
This is why we think there is some problem with trip linking.
```{r}
ggplot(data=trip_trace_quality_data_rulesy, aes(x=error_flag, y=trip_rulesy, fill= final_label))+

  geom_bar(stat="identity")+ theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
  
```


```{r}
trip_raw%>%group_by(user_added,diary_platform)%>%count()
trip_raw%>%group_by(user_split,diary_platform)%>%count()
trip_raw%>%group_by(user_merged,diary_platform)%>%count()

trip_raw%>%group_by(diary_platform)%>%count()

```



#Summaries of change mode purpose trips only, post-rulesy

```{r}

example_recs_mode_1<-trips_edits%>%filter(error_flag=='mode_1 missing')

#example_recs_missing_trip_link<-trip_rulesy_raw%>%filter(error_flag=='missing next trip link')
```

### function to summarize avariable across a single flag
```{r}
var_compare<-function(df, id, var_name){
  #id=!!ensym(id)
  #var_name=!!ensym(var_name)

  df_cols<-df%>%select(!!ensym(id), !!ensym(var_name))
  vals<- values%>%filter(variable==sym(var_name))
  df_cols<-df_cols%>%
    left_join(vals, by=join_by(!!ensym(var_name)==value))
  df_summary<-df_cols%>%
              group_by(final_label)%>%
              count()

  
}
```


```{r}

trip_rulesy_col<-trip_rulesy%>%
  select(hhid,error_flag, person_id, day_id, tripnum, tripid, depart_time_timestamp, arrival_time_timestamp, mode_1, distance_miles, travel_time, speed_mph, origin_lat, origin_lng, dest_lat, dest_lng, origin_purpose,dest_purpose,depart_time_timestamp,arrival_time_timestamp, arrive_date, arrival_time_hour, arrival_time_minute, depart_time_hour, depart_time_minute)%>%arrange(hhid,person_id, depart_time_timestamp, arrival_time_timestamp)

```

```{r}
trip_rulesy_hh<-trip_rulesy_col%>%filter(hhid==23014826)

```

