---
title: Summarizing Rulesy Error Codes
  Fixie
author: "PSRC"
date: "2023-10-04"
output:
  pdf_document:
    toc: yes
    toc_depth: '4'
  html_document:
    toc: yes
    toc_depth: 4
    toc_float: yes
    css: styles.css
---

```{r }
library(data.table) # general data manipulation
library(srvyr)      # survey statistics
library(dplyr)      # helpers for srvyr (mutate, group_by, and %>% pipe), 
                    # plus function near(), used in qa/qc checks 
library(stringr)    # string helper functions
library(flextable)  # formatted tables
library(ggplot2)    # plotting
library(scales)     # ggplot2 helpers
library(plotly)     # interactive plots
library(rstudioapi) # for setting working directory
library(psrcelmer)
library(tidyr)

# Point to where the 2023 codebook lives:
codebook_path = 'J:/Projects/Surveys/HHTravel/Survey2023/Data/data_deliverable_81823/codebook_guide/PSRC_Combined_Codebook_2023_08162023.xlsx'

# Read Data  ===================================================================



hh_rulesy<-get_table(db_name = "HouseholdTravelSurvey2023", "combined_data", "v_household")
person_rulesy<-get_table(db_name = "HouseholdTravelSurvey2023", "combined_data", "v_person")
day_rulesy<-get_table(db_name = "HouseholdTravelSurvey2023", "combined_data", "v_day")
trip_raw<-get_table(db_name = "HouseholdTravelSurvey2023", "combined_data", "v_trip")
trip_rulesy_before<-get_table(db_name = "hhts_cleaning", "HHSurvey", "Trip")
trip_error_flags<-get_table(db_name = "hhts_cleaning", "HHSurvey", "trip_error_flags")
error_types<-get_table(db_name = "hhts_cleaning", "HHSurvey", "error_types")





## Codebook -------------------------------------------------------------------- 

variables = readxl::read_xlsx(
  codebook_path,
  sheet = 'variable_list_2023'
)

values =  readxl::read_xlsx(
  codebook_path,
  sheet = 'value_labels_2023'
)

setDT(variables)
setDT(values)
```

Create visualizable data tables, before and after the changes. Also show the errors
```{r}
trip_raw<-trip_raw%>%mutate(tripid=as.character(tripid))
trip_rulesy<-trip_rulesy_before%>%mutate(tripid=as.character(tripid))
trip_rulesy<-left_join(trip_rulesy, trip_error_flags, by = c('person_id', 'tripnum'), suffix=c('_rulesy', '_error'))%>%group_by(person_id, tripnum)%>%filter(row_number()==1)
#trip_rulesy<-trip_rulesy%>%mutate(person_id=as.character(person_id))
#person_rulesy<-person_rulesy%>%mutate(person_id=as.character(person_id))
#trip_rulesy<-left_join(trip_rulesy, person_rulesy, by ='person_id')

trip_rulesy_subset<-trip_rulesy%>%
  select(error_flag, person_id, day_id, tripnum, tripid, depart_time_timestamp, arrival_time_timestamp, mode_1, distance_miles, travel_time, speed_mph, origin_lat, origin_lng, dest_lat, dest_lng, origin_purpose,dest_purpose,depart_time_timestamp,arrival_time_timestamp, arrive_date, arrival_time_hour, arrival_time_minute, depart_time_hour, depart_time_minute)%>%arrange(person_id, depart_time_timestamp, arrival_time_timestamp)%>%head(500)

trip_rulesy_raw<-left_join(trip_raw, trip_rulesy, by='tripid', suffix=c('_before', '_after'))%>%
  select(error_flag,person_id_before, day_id_before, day_id_after, tripnum_after, tripid,origin_purpose_before, origin_purpose_after, dest_purpose_before, dest_purpose_after, arrive_date_before, arrive_date_after,arrival_time_hour_before,arrival_time_minute_before,depart_time_hour_before, depart_time_minute_before, depart_time_timestamp, arrival_time_timestamp,depart_time_minute_after,origin_lng_before, origin_lat_before,dest_lng_before,dest_lng_after, dest_lat_before,dest_lat_after, mode_1_before, mode_1_after, trace_quality_flag_before, day_is_complete_a_before, day_is_complete_b_before)%>%arrange(person_id_before, arrive_date_before, arrival_time_hour_before, arrival_time_minute_before, depart_time_hour_before, depart_time_minute_before)%>%head(500)


```


select all records of a person who has a particular error mode
```{r}
# first find the person ids of  trips with the error code

##trip_rulesy%>%s


```



### function to summarize a variable across flags
```{r}
var_error_compare<-function(df, dfname, id, var_name){
  #id=!!ensym(id)
  #var_name=!!ensym(var_name)
  df<-df%>%filter(!is.na(error_flag))
  df_cols<-df%>%select(!!ensym(id), !!ensym(var_name), error_flag)
  vals<- values%>%filter(variable==sym(var_name))
  df_cols<-df_cols%>%
    left_join(vals, by=join_by(!!ensym(var_name)==value))
  df_summary<-df_cols%>%
              group_by(final_label, error_flag)%>%
              count()
  df_summary<-df_summary%>%
              rename(!!quo_name(dfname):=n)
  
}
```


```{r}

trip_mode_data_rulesy<-var_error_compare(df=trip_rulesy,dfname='trip_rulesy',id='tripid', var_name='mode_type')

```

### Error flags by mode
There are very high numbers of walk, transit trips with the change mode purpose and/or missing next trip link or missing prior trip link.
This is why we think there is some problem with trip linking.
```{r}
ggplot(data=trip_mode_data_rulesy, aes(x=error_flag, y=trip_rulesy, fill= final_label))+

  geom_bar(stat="identity")+ theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
  
```




#Summaries of change mode purpose trips only, post-rulesy

```{r}

example_recs_change_mode<-trip_rulesy_raw%>%filter(error_flag=='"change mode" purpose')

example_recs_missing_trip_link<-trip_rulesy_raw%>%filter(error_flag=='missing next trip link')
```

### function to summarize avariable across a single flag
```{r}
var_compare<-function(df, id, var_name){
  #id=!!ensym(id)
  #var_name=!!ensym(var_name)

  df_cols<-df%>%select(!!ensym(id), !!ensym(var_name))
  vals<- values%>%filter(variable==sym(var_name))
  df_cols<-df_cols%>%
    left_join(vals, by=join_by(!!ensym(var_name)==value))
  df_summary<-df_cols%>%
              group_by(final_label)%>%
              count()

  
}
```


